---
title: "Assignment 2"
author: "Dushyant Goel"
date: "2024-09-25"
output: html_document # you can change to other output format if you want
---

```{r}
library(tidyverse)
library(Stat2Data)
data("Hawks")
```

# 1. Data Wrangling

## 1.1 (Q1)

```{r}
hSF <- Hawks %>%
filter(Species == "RT", Weight >= 1000) %>%
select(Wing, Weight, Tail) 
```

## 1.1 (Q2)

The data has `3` variables. In Machine Learning parlance, there are `3` features. There are `398` observables, which in different context may be called "examples" or "observations"

## 1.2 (Q1)

```{r}
hSF %>%
arrange(Wing) %>%
head(n=5)
```

## 1.3 (Q1)

```{r}
hawkSpeciesNameCodes <-
data.frame(species_code = c('CH', 'RT', 'SS'), species_name_full = c("Cooper's", "Red-tailed", "Sharp-shinned"))
```

## 1.3 (Q2)

```{r}
hawksFullName <- Hawks %>%
left_join(hawkSpeciesNameCodes, by=c("Species"="species_code")) %>%
select(-Species) %>%
rename(Species = species_name_full)
```

## 1.3 (Q3)

```{r}
hawksFullName %>% select(Species, Wing, Weight) %>% head(n=7)
```

Yes, it matters what type of join we use. There are 4 options:

1.  Inner Join -\> matching column must be present in both tables, else the row is ignored.

2.  Left Join -\> All rows of left table are retained and matching rows from right tables are appended.

3.  Right Join -\> All rows of right table are retained and matching rows from left tables are appended.

4.  Outer Join -\> Only rows of both tables are retained.

It would have made a difference if the column `Species` in the data.frame `Hawks` had NA values or values different from "CH", "RT" or "SS".

### 1.4 (Q1)

```{r}
hawksWithBMI <- Hawks %>%
mutate(bird_BMI = (Weight*1000/((Wing)**2))) %>%
arrange(desc(bird_BMI)) %>%
select(Species, bird_BMI)
```

NB: To match the values of BMI in reference, the original weight must be assumed in `kgs` and wingspan in `mm`. It is comical then, to think of haws weighing a ton but flying with wings of less than 500 mm.

## 1.5 (Q1)

```{r}
hawksFullName %>%
group_by(Species) %>%
summarise(num_rows = n(), mn_wing = mean(Wing, na.rm=TRUE), md_wing = median(Wing, na.rm=TRUE), t_mn_wing = mean(Wing, trim=0.1, na.rm=TRUE), b_wt_ration = max(Wing/Tail, na.rm=TRUE))

```

## 1.5 (Q2)

```{r}
hawksFullName %>%
select(Wing, Weight, Culmen, Hallux, Tail, StandardTail, Tarsus, Species) %>%
group_by(Species) %>%
summarise(across(everything(), ~sum(is.na(.x))))
```

# 2. Random experiments, events and sample spaces, and the set theory

## 2.1 (Q1)

**Random Experiment** : A random experiment is an event which - has a well defined set of outcomes. - can be repeated infinitely many times.

## 2.1 (Q2)

When we roll a dice, this is a random experiment. The sample space\
$\Omega$ = {1, 2, 3, 4, 5, 6}

An event, $\epsilon$ is any (or none) combination of events from $\Omega$

For example, $\epsilon = \{1,2,3\}$.\

Since we can pick none or any number of elements of $\Omega$, it follows that the total number of events is the cardinality of superset of $\Omega$. Since $\Omega$ contains $6$ elements, the number of events possible is\
$2^6 = 64$\

By definition, the empty set $\phi$ is also considered an event.

## 2.2 (Q1)

$$

A := \{1, 2, 3\} \\
B := \{2, 4, 6\} \\
C := \{4, 5, 6\} \\

$$

1.  $ A \cup B := \{1, 2, 3, 4, 6\} $ 
2.  $ A \cap B := \{2\} $ , $ A \cap C := \phi $ 
3.  $ A \setminus B := \{1, 3\}$, $ A \setminus C := \{1, 2, 3\} $ 
4.  $ A $ and $ B $ are not disjoint, $ A $ and $ C $ are disjoint.
5.  By definition of complementary, $ B $ and $ A :setminus B $ are disjoint sets.
6.  $ {1, 2, 3, 4, 5, 6 } $\
    Partition into $two$ sets, say even and odd. $\{\{1, 3, 5\}, \{2, 4, 6\}\}$\
    Partition into $three$ sets, say $\{\{1, 4\}, \{2, 5\}, \{3, 6\}\}$\

## 2.2 (Q2)

1.  $$
    
    (A^c)^c  =>\\
    \omega \in \Omega : \omega \notin A^c =>\\
    \omega \in A => \\
    A
    
    $$
2.  $\phi$
3.  Suppose $ x \in B^c $\
    $ => x \notin B $ --- 1\
    $ A \subseteq B $\
    $ => \forall y \in B, y \in A $\

Combining 1 and 2 $ if y \notin B => y \notin A $ --- 2\
$ y \in B^c => y \in A^c $\
$ B^c \subseteq A^c $

4.  To Prove:\
    $ (A \cap B)^c = A^c \cup B^c $\

Suppose,\
$ x \in (A \cap B)^c $\
$ x \notin (A \cap B) $\
$ x \ is \ not \ in \ A \ and \ x \ is \ not \ in \ B $\
$ x \notin A \ and \ x \notin B $ \
$ x \in A^c and x \in B^c $ \
$ x \in A^c \cap B^c $ \

Expression for $(\bigcup_{k=1}^{K} A_k)^c $ \
$\bigcap_{k=1}^{K} A_k ^c $\

5. To Prove : \$ (A \cup B)\^c = A\^c \cap B\^c \$

Suppose,\
\$ x \in (A \cup B)\^c \$\
\$ x \notin (A \cup B) \$\
\$ x \notin A and x \notin B \$\
\$ x \notin A\^c and x \notin B\^c $\$ x \in A\^c \cap B\^c \$

6. Expression for $(\bigcap_{k=1}^{K} A_k)^c $ \
$\bigcup_{k=1}^{K} A_k ^c $

## 2.2 (Q3)

$ |E| := 2^K - 1 $
This is because every element in $\omega$ has $2$ options, 
- to be chosen as an element of a set, which is contained in E.
- not be chosen.

But Since $ E \subset \omega$ and not $ E \subseteq \omega$, we substract the case when all elements of $\omega$ are chosen.

## 2.2 (Q4)

1. $\phi$
2. 
- $ S_1, S_2, S_3, S_4 $ are pairwise disjoint.
- $ S_1 \cup S_2 \cup S_3 \cup S_4 = A_1 \cup A_2 \cup A_3 \cup A_4 $

Hence the given $ \bigcup S{i} $ form a partition.

# 2.2 (Q5)
1. $\1_{A^c(}(\omega) := 1 - 1_A(\omega)$
2. $\Omega$
3. $ 1_{(A \cap B)^c} => $ \
$ 1 - 1_{A \cap B} => $ \
$ 1 - 1_A.1_B => $ \
$ 1 - 1_A + 1 - 1_B - 1 + 1_A + 1_B - 1_A.1_B $ (adding and subtracting $1_A, 1_B$ \
$ (1-1_A) + (1-1_B) - (1-1_A)(1-1_B) $ \
$ (1_{A^c}) + 1_{B^c} - 1_{A^c}.1_{B^c} $
$ 1_{A^c \cup B^c} $

$ Q.E.D. $


# 3. Probability theory

## 3.1 (Q1)

$ \epsilon = \{ {\phi\}, \{a\}, \{b\}, \{c\}, \{a,b\}, \{b,c\}, \{a,c\}, \{a,b,c\}  \} $ \

We define $ P(A_i) \forall A_i \in \epsilon. $ Here $ A_i := \{a\} , \{b\}, \{c\}$ are disjoint events. \  

P{$\phi$} = 0 \
P({a}) = 0.5 \
P({b}) = 0.1 \
P({c}) = 0.4 \
P({a,b}) = 0.6 \
P({b,c}) = 0.5 \
P({a,c}) = 0.9 \
P({a,b,c}) = 1.0 \


## 3.1 (Q2)

The given function $P$ will satisfie the Rules of Probability if the disjoint  paritions {0} and {1} have probability $\geq$ 0 and their sum is 1.

However, that depends on value of $q$. \
If $ 0 \leq q \leq 1 $, then the given function is a probability function.

## 3.2 (Q1)
We know, for sets $ A $ and $ B $, \
$ P(A \cup B) = P(A) + P(B) - P(A \cap B) $ \ 
For disjoint sets $P(A \cap B) = 0$


Consider $P(\bigcup_{i=1}^n A_i)$,  will be the sum of $P(A_i)$ and subtracting pairwise $P(A_i \cap A_{i+1})$. But we know for disjoint $A_i$ all $P(A_i \cap A_{i+1})$ = 0. \

Hence the intersection has no contribution to the union. \

Using **Rule 3**, \

$P(\bigcup_{i=1}^n A_i) = \sum_{i=1}^nP(A_i)$

## 3.2 (Q2)

$$
P(S^c) \\
= P(\Omega \setminus S) \\
= P(\Omega) - P(S) \\
= 1 - P(S) \\
$$
Q.E.D.

## 3.3 (Q3)

Let the sample space $\Omega$ be the set of all natural numbers $N$. \
Let us define an experiment, in which we pick up one number with replacement from $N$ and divide it by 2 and by 5.\
Let the events $\omega_1$ := number is divisble by 2, $\omega_2$ := number is divisible by 5. \
$P({\omega_1}) = \frac{1}{2}$ -- a \
$P({\omega_2}) = \frac{1}{5}$ -- b \
$P({\omega_1, \omega_2}) = \frac{1}{2} + \frac{1}{5} - \frac{1}{10}$ -- c \

$ a + b \neq c $ \
We can generalize this to $\omega_i$ event, where the $i^{th}$ event represents if the number is divisible by i+1. \

Hence proven.

## 3.3 (Q4)

The probabilities of disjoint sets is simple summation. Breaking the problem into disjoint sets :-
$$
P(A \cup B) = P(A\cap B^c) + P(A\cap B) + P(A^c \cap B) -- 1
$$
Now
$$
P(A) = P(A \cap B^c) + P(A \cap B) -- 2 \\
P(B) = P(B \cap A^c) + P(A \cap B) -- 3 \\
$$

Rearranging 2, 3 and substituting $P(A\cap B^c), P(A^c \cap B)$ values in 1

$$
P(A \cup B) = P(A) - P(A \cap B) + P(B) - P(A \cap B) + P(A \cap B) \\
=> P(A \cap B) = P(A) + P(B) - P(A \cap B)
$$

Hence Proven