---
title: "Assignment 2"
author: "Dushyant Goelli"
date: "2024-09-25"
output: html_document # you can change to other output format if you want
---

# 1. Data Wrangling

## 1.1 (Q1)

```{r}
hSF <- Hawks %>%
filter(Species == "RT", Weight >= 1000) %>%
select(Wing, Weight, Tail) 
```

## 1.1 (Q2)

The data has `3` variables. In Machine Learning parlance, there are `3` features. There are `398` observables, which in different context may be called "examples" or "observations"

## 1.2 (Q1)

```{r}
hSF %>%
arrange(Wing) %>%
head(n=5)
```

## 1.3 (Q1)

```{r}
hawkSpeciesNameCodes <-
data.frame(species_code = c('CH', 'RT', 'SS'), species_name_full = c("Cooper's", "Red-tailed", "Sharp-shinned"))
```

## 1.3 (Q2)

```{r}
Hawks %>%
left_join(hawkSpeciesNameCodes, by=c("Species"="species_code")) %>%
select(-Species) %>%
rename(Species = species_name_full)
```

## 1.3 (Q3)

```{r}
hawksFullName %>% select(Species, Wing, Weight) %>% head(n=7)
```

Yes, it matters what type of join we use. There are 4 options: 1. Inner Join -\> matching column must be present in both tables, else the row is ignored. 2. Left Join -\> All rows of left table are retained and matching rows from right tables are appended. 3. Right Join -\> All rows of right table are retained and matching rows from left tables are appended. 4. Outer Join -\> Only rows of both tables are retained.

It would have made a differece if the column `Species` in the data.frame `Hawks` had NA values or values different from "CH", "RT" or "SS".

### 1.4 (Q1)

```{r}
hawksWithBMI <- Hawks %>%
mutate(bird_BMI = (Weight*1000/((Wing)**2))) %>%
arrange(desc(bird_BMI)) %>%
select(Species, bird_BMI)
```

"""

NB: To match the values of BMI in reference, the original weight must be assumed in `kgs` and wingspan in `mm`. It is comical then, to think of haws weighing a ton but flying with wings of less than 500 mm.

"""

## 1.5 (Q1)

```{r}
hawksFullName %>%
group_by(Species) %>%
summarise(num_rows = n(), mn_wing = mean(Wing, na.rm=TRUE), md_wing = median(Wing, na.rm=TRUE), t_mn_wing = mean(Wing, trim=0.1, na.rm=TRUE), b_wt_ration = max(Wing/Tail, na.rm=TRUE))

```

## 1.5 (Q2)

```{r}
hawksFullName %>%
select(Wing, Weight, Culmen, Hallux, Tail, StandardTail, Tarsus, Species) %>%
group_by(Species) %>%
summarise(across(everything(), ~sum(is.na(.x))))
```


# 2. Random experiments, events and sample spaces, and the set theory

....

# 3. Probability theory

## 3 (Q1)

## 3 (Q2)

.....
