---
title: "Assignment 2"
author: "Dushyant Goel"
date: "2024-09-25"
output: pdf_document # you can change to other output format if you want
---

```{r}
library(tidyverse)
library(Stat2Data)
data("Hawks")
```

# 1. Data Wrangling

## 1.1 Select and Filter
### (Q1)

```{r}
hsf <- Hawks %>%
  filter(Species == "RT", Weight >= 1000) %>%
  select(Wing, Weight, Tail)

head(hsf, 5)
```

### (Q2)

The data has `3` variables. In Machine Learning parlance, there are `3` features.   
There are `398` observables, which in different context may be called "examples" or "observations"  

## 1.2 The arrange function 
### (Q1)

```{r}
hsf %>%
  arrange(Wing) %>%
  head(n = 5)
```

## 1.3 Join and rename functions
### (Q1)

```{r}
species_code <- c("CH", "RT", "SS")
species_name_full <- c("Cooper's", "Red-tailed", "Sharp-shinned")
hawk_species_name_codes <-
  data.frame(species_code, species_name_full)

hawk_species_name_codes
```

### (Q2)

```{r}
hawks_full_name <- Hawks %>%
  left_join(hawk_species_name_codes, by = c("Species" = "species_code")) %>%
  select(-Species) %>%
  rename(Species = species_name_full)
```

### (Q3)

```{r}
hawks_full_name %>%
  select(Species, Wing, Weight) %>%
  head(n = 7)
```

Yes, it matters what type of join we use. There are 4 options:

1.  Inner Join -\> matching column must be present in both tables, else the row is ignored.

2.  Left Join -\> All rows of left table are retained and matching rows from right tables are appended.

3.  Right Join -\> All rows of right table are retained and matching rows from left tables are appended.

4.  Outer Join -\> Only rows of both tables are retained.

It would have made a difference if the column `Species` in the data.frame `Hawks` had NA values or values different from "CH", "RT" or "SS".

## 1.4 The mutate function 

### (Q1)

```{r}
hawks_with_bmi <- Hawks %>%
  mutate(bird_BMI = (Weight * 1000 / ((Wing)**2))) %>%
  select(Species, bird_BMI) %>%
  arrange(desc(bird_BMI))

hawks_with_bmi %>%
  head(8)
```

NB: To match the values of BMI in reference, the original weight must be assumed in `kgs` and wingspan in `mm`. It is comical then, to think of hawks weighing a ton but flying with wings of less than 500 mm.

## 1.5 Summarize and group-by functions
### (Q1)

```{r}
hawks_full_name %>%
  group_by(Species) %>%
  summarise(num_rows = n(),
            mn_wing = mean(Wing, na.rm = TRUE),
            md_wing = median(Wing, na.rm = TRUE),
            t_mn_wing = mean(Wing, trim = 0.1, na.rm = TRUE),
            b_wt_ration = max(Wing / Tail, na.rm = TRUE))

```

### (Q2)

```{r}
hawks_full_name %>%
  select(Wing, Weight, Culmen, Hallux, Tail, StandardTail, Tarsus, Species) %>%
  group_by(Species) %>%
  summarise(across(everything(), ~sum(is.na(.x))))
```

---


# 2. Random experiments, events and sample spaces, and the set theory
## Random experiments, events and sample space
### (Q1)

**Random Experiment** : A random experiment is an event which 
- has a well defined set of outcomes. 
- can be repeated infinitely many times.

### (Q2)

When we roll a dice twice, this is a random experiment. The sample space\
$\Omega$ = {(a,b) | a,b $\in$ {1, 2, 3, 4, 5, 6}}

There are $36$ possible outcomes.\
Since we can pick none or any number of elements of $\Omega$, it follows that the total number of events is the cardinality of superset of $\Omega$. Since $\Omega$ contains $36$ elements, the number of events possible is\
$2^{36}$\

By definition, the empty set $\phi$ is also considered an event.


## 2.2 Set Theory 
### (Q1) Set Operations

$A := {1, 2, 3} \\ B := {2, 4, 6} \\ C := {4, 5, 6}$

1.  $A \cup B := {1, 2, 3, 4, 6}$
2.  $A \cap B := {2} \\ A \cap C := \phi$
3.  $A \setminus B := \{1, 3\} \\ A \setminus C := \{1, 2, 3\}$
4.  $A$ and $B$ are not disjoint, $A$ and $C$ are disjoint.
5.  By definition of complementary, $B$ and $A \setminus B$ are disjoint sets.
6.  $\{1, 2, 3, 4, 5, 6 \}$  
    Partition into $two$ sets, say even and odd ${\{1, 3, 5\}, \{2, 4, 6\}}$  
    Partition into $three$ sets, say ${\{1, 4\}, \{2, 5\}, \{3, 6\}}$  

### (Q2) Complements, subsets and De Morgan's Law

1.  $A$
2.  $\phi$
3.  Suppose $x \in B^c$  
$\implies x \notin B$  
$\implies x \notin A$  
$\implies x \in A^c$  
This holds $\forall x \in B$, hence $A^c \subseteq B^c$
4.  To Prove:  
    $(A \cap B)^c = A^c \cup B^c$  
Part 1 : $x \in (A \cap B)^c \implies x \in A^c \cup B^c$  
Let $x \in (A \cap B)^c$  
Then, $x \notin A \cap B$  
Then, either $x \notin A$ or $x \notin B$ or both  
if $x \notin A$ then $x \in A^c$  
if $x \notin B$ then $x \in B^c$  
in either case, $x \in A^c \cup B^c$  
Part 2: $x \in A^c \cup B^c \implies x \in (A \cap B)^c$  
Let $x \in A^c \cup B^c$  
Assume $x \notin (A \cap B)^c$  
Then, $x \in A \cap B$  
Then, $x \in A\ and\ x \in B$  
Then, $x \notin A^c\ and\ x \notin B^c$  
Then, $x \notin A^c \cup B^c$  
This is a contradiction, hence $x \in A^c \cup B^c$  
From part 1 and part 2: $(A \cap B)^c = A^c \cup B^c$  
Hence proven.  
$(\bigcap_{k=1}^{K}A_k)^c = \bigcup_{k=1}^{K}(A_k)^c$  
5.  To Prove:  
$(A \cup B)^c = A^c \cap B^c$  
Part 1: $x \in (A \cup B)^c \implies x \in A^c \cap B^c$  
Let $x \in (A \cup B)^c$  
Then, $x \notin A \cup B$  
Then, $x \notin A\ and\ x \notin B$  
Then, $x \in A^c\ and\ x \in B^c$
Then, $x \in A^c \cap B^c$  
Part 2: $x \in A^c \cap B^c \implies x \in (A \cup B)^c$  
Let, $x \in A^c \cap B^c$  
Assume, $x \notin (A \cup B)^c$  
Then, $x \in (A \cup B)^c$  
Either, $x \in A$ or $x \in B$ or both  
In both cases, $x \notin A^c \cap B^c$  
This is a contradiction, hence $x \in A^c \cup B^c$  
From part 1 and part 2: $(A \cup B)^c = A^c \cap B^c$  
Hence proven.

6. $(\bigcap_{k=1}^{K} A_k)^c = \bigcup_{k=1}^{K} A_k^c$  

### (Q3)

$\|E\| := 2^K - 1$  
It is not $2^K$ because the set $\Omega$ is not included in $E$ as defined. 

### (Q4)

1.  $\phi$
2.  To Prove:
- $S_1 \cup S_2 \cup S_3 \cup S_4 = A_1 \cup A_2 \cup A_3 \cup A_4$  
Part 1: Let $x \in S_1 \cup S_2 \cup S_3 \cup S_4$  
$\forall i\ if\ x \in S_i$   
$\implies x \in A_i \setminus _{j < i}A_j$  
$\implies x \in A_i$  
Part 2: Let $x \in A_1 \cup A_2 \cup A_3 \cup A_4$  
$\forall i\ if\ x \in A_i$  
$\implies x \in A_i \setminus _{j < i} A_i$ for sufficiently small $i$  
$\implies x \in S_i$  
From part 1 and part 2 $S_1 \cup S_2 \cup S_3 \cup S_4 = A_1 \cup A_2 \cup A_3 \cup A_4$  
- $S_1, S_2, S_3, S_4$ are pairwise disjoint.  
Suppose $x \in S_{i_0} \cap S_{i_1}$ for some $i_0 < i_1$  
Then, $x \in S_{i_1}$  
$\implies x \in A_{i_1} \setminus _{i < i_1} A_i$  
$\implies x \notin A_{i_0}$ (as $i_0 < i_1$)  
$\implies x \notin A_{i_0} \setminus _{i < i_0} A_i$  
$\implies x \notin S_{i_0}$  
This is a contradiction, hence there is no $x$ such that $x \in S_{i_0} \cap S_{i_1}$  
i.e. $S_{i_0} and S_{i_1}$ are pairwise disjoint

Hence the given $\bigcup S_{i}$ form a partition.

# 3. Probability theory

## 3.1 Rules of probability
### (Q1)

$E = \{ {\phi}, \{a\}, \{b\}, \{c\}, \{a,b\}, \{b,c\}, \{a,c\}, \{a,b,c\} \}$
$$
P(\phi) = 0\\
P(\{a\}) = 0.5\\
P(\{b\}) = 0.1\\
P(\{c\}) = 0.4\\
P(\{a,b\}) = 0.6\\
P(\{b,c\}) = 0.5\\
P(\{a,c\}) = 0.9\\
P(\{a,b,c\}) = 1.0\\
$$

### (Q2)

Rule 1: $P(A) \geq 0$ for any $A \in E$  
Given $$
P(\phi) = 0 \\
P(\{0\}) = 1-q \\
P(\{1\}) = q \\
P(\{0,1\}) = 1 \\
$$
This satisfies Rule 1

Rule 2: $P(\Omega) = 1$  
Here $\Omega = \{0, 1\}$. Given $P(\{0,1\}) = 1$  
This satisfies Rule 2

Rule 3: Pairwise disjoint events are $\{0\}, \{1\}$.  
$P(\{0\} \cup \{1\}) = P(\{0, 1\}) = 1$  
$P(\{0\}) + P(\{1\}) = q + (1-q) = 1$  
LHS = RHS  
This satisfies Rule 3  

## 3.2 Deriving new properties from the rules of probability 
### (Q1)


Using **Rule 3**,\
$P(\bigcup _{i=1}^{\infty} A_i) = \sum_{i=1}^\infty P(A_i)$  
Define $A_{n+1}=\phi, A_{n+2}=\phi, ...$  
$P(\phi) = 0$

Now apply Rule 3 on given equation to get the result.

### (Q2)
By Rule 3,
$$
P(S^c) + P(S) = P(S \cup S^c) \\
P(S^c) + P(S) = P(\Omega) \\
P(S^c) + P(S) = 1 ...[rule\ 2]\\ 
P(S^c) = 1 - P(S) \\
$$ Q.E.D.

### (Q3)

Let the sample space $\Omega$ be the set of all natural numbers $N$.  
Let the events $S_1$ := number is divisble by 2, $S_2$ := number is divisible by 5.  
Then,  
$P({S_1}) = \frac{1}{2}$ \
$P({S_2}) = \frac{1}{5}$ \
$P({S_1 \cup S_2}) = \frac{1}{2} + \frac{1}{5} - \frac{1}{10} = \frac{3}{5}$ \
$P(S_1) + P(S_2) = \frac{1}{2} + \frac{1}{5} = \frac{7}{10}$  
$1 \neq 2$  
We can generalize this to $S_i$ event, where the $i^{th}$ event represents if the number is divisible by i+1.\

Hence proven.

### (Q4)

Using Rule 3,
$$
P(A \cup B) = P(B) + P(B \setminus A) ..1 \\
P(A \cup B) = P(A) + P(A \setminus B) ..2 \\
P(A \cup B) = P(B \setminus A) + P(A \setminus B) + P(A \cap B) ..3 \\
$$

1 + 2 - 3 =>

$$
P(A \cup B) = P(A) + P(B) - P(A \cap B)
$$
Hence Proven
